---
title: "lecture_notes09_13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
glimpse(diamonds)
```

## Concept 1
Split - Apply - Combine Concept
[group]

## Concept 2
Piping %>%
feeding the return value to the first argument of next function
allows for better "reading" of R code as sentences (more of a flow chart than disconnected code)
take this data frame, group the data into unique values, THEN calculate things and make a plot for example

what mappings would we use ? geom - point\

Functions that return ONE result PER GROUP = summarise()

Functions that return ONE result PER ROW = mutate()

```{r}
ggplot(diamonds, aes(x = clarity,y = price))
  geom_boxplot()
```

lets calculate the average (and std. deviation) for price for each clarity category

```{r}
clarity_cost <- diamonds %>%
  group_by(clarity) %>% #split
  summarise(average.cost = mean(price, na.rm = TRUE), #function in R for average is mean! na.rm = ignore N/A
            std.dev = sd(price, na.rm =TRUE)) 
  ggplot(clarity_cost, aes(x=clarity,y=average.cost)) + 
    geom_pointrange(aes(ymin = average.cost - std.dev,
                        ymax = average.cost + std.dev))
```

## Concept 3 - mutate

```{r}
price_per_carat <- diamonds %>%
  # group_by(clarity_cost) %>%
  mutate(cost.per.carat = price / carat)

ggplot(price_per_carat, aes(x = cost.per.carat, color =  clarity)) + 
  geom_density()
```




